Peaks Over Threshold for Bursty Time Series

### 2018-04-18 article_springer/article_springer.Rmd
write conclusion & appendix

  This article generalizes the Peaks Over Threshold (POT) model to the setting 
  where inter-event times are heavy-tailed and i.i.d.. For high thresholds and 
  infinite-mean waiting times, we show that the times between threshold 
  crossings are Mittag-Leffler distributed, and thus form a "fractional 
  Poisson Process" which generalizes the standard Poisson Process.
  We model threshold crossing times and threshold exceedances jointly, and
  provide graphical means of estimating model parameters and assessing 
  model fit. Finally, we study two real-world, bursty time series, and 
  show how to extract parameter estimates as well as estimates for the time 
  of the next threshold crossing.

@@

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Sections 3 &
4. 
A statistical procedure to estimate model parameters via stability plots 
is outlined in Section 5, and diagnostic plots for model criticism are 
discussed in Section 6. 
In Section 7, we give case studies for simulated data as well as for a
real-world dataset, with a discussion in Section 8. 
All statistical computations have been done in R, and all code, data, as 
well as the source code for this manuscript has been organized into an R 
software package, available at <https://github.com/strakaps/CTRM>.

@@

: Let the waiting times $J_k$ be in the domain of attraction of a 
positively skewed sum-stable law with stability parameter $0 < \beta < 1$; 
more precisely,
$$
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
$$
for a function $b(n)$ which is regularly varying at $\infty$ with 
parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$. 
Write $p := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p)} \to W_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $W_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sW_\beta)] = \frac{1}{1+s^\beta}.
$$

@@

@@

```{r solar-flare-tail-scale, message=FALSE, fig.height=4.5, fig.width=8, fig.cap="\\label{fig:flares}Stability plots for the tail and scale parameter of the Mittag-Leffler distribution of the Solar Flare dataset. Dotted horizontal lines are at $\\beta = 0.85$ and $\\sigma_0 = 3 \\times 10^7$."}
thin_flares_ctrm <- flares %>% ctrm() %>% thin(k = 700)
par(mfrow=c(1,2))
MLestimates(thin_flares_ctrm, tail = 0.9, scale = 3 * 1E7)
```

@@

# Application to data

## Simulated Data

To test our inference method, we have simulated `r params$n` 
independent
waiting time and magnitude pairs $(W_k, J_k)$. In order to have exact 
analytical values available for $\beta$ and $\sigma_0$, a distribution for 
$W_k$ needs to be chosen for which $b(n)$ from \eqref{eq:Wscale} is known. 
If we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{theD}, 
then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation (see e.g. the vignette on parametrisation in @MittagLeffleR) 
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist.

@@

By \eqref{eq:Tellscale}, the distribution 
of $T(\ell)$ is approximately 
${\rm ML}(\beta, p^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta})$, which means 
$\sigma_0 = n^{1/\beta}$.
The distribution of $J_k$ is irrelevant for the inference on $\beta$ and 
$\sigma_0$ (we have chosen unit exponential random variables). 
Figure \ref{fig:sim} displays plots of $\hat \beta(\ell)$ and 
$\hat \sigma(\ell)$ vs. $k$, the number of exceedances as $\ell$ traverses 
the largest observations $J_i$. 
Dotted lines show 95% confidence intervals, which are derived from the 
asymptotic normality of the log-moments estimators [@Cahoy2013] and the $\delta$-method [@MittagLeffleR].
The dashed lines show the actual values of $\beta$ resp.\ $\sigma_0$, 
showing that our inference method effectively identifies the 
right parameters.

## Bitcoin trading volumes

```{r bitcoin-data, fig.cap="\\label{fig:bitcoin-data}Time Series of 'Bitcoin days destroyed', a measure for the traded volume of bitcoins."}
bitcoin_ctrm <- ctrm(bitcoin)
plot(bitcoin_ctrm, p = 0.02)
```

We study the daily trade volumes of Bitcoin since 2009, provided by 
Blockchain and downloaded from Data 
Market\footnote{\url{https://datamarket.com/data/list/?q=provider:blockchain}}. 
Although the observations are regular (daily), extremes in the trade volumes
occur in bursts, which can be seen over timescales of months and years 
(Figure \ref{fig:bitcoin-data}).

```{r bitcoin-data-tail-scale, message=FALSE, fig.cap="\\label{fig:bitcoin-tail-scale}Tail and scale plot of the bitcoin data."}
thin_bitcoin_ctrm <- thin(ctrm = bitcoin_ctrm, k = 250)
par(mfrow=c(1,2))
MLestimates(thin_bitcoin_ctrm, tail = 0.9, scale = 2 * 1E3)
```

Figure \ref{fig:bitcoin-tail-scale} shows plots identifying the candidates
$\beta = 0.9$ and $\sigma_0 = 2000$ for 
the tail and scale parameters of the exceedance time distribution. 
That is, exceedances of a threshold $\ell$ at the $k$-th largest observed 
magnitude are distributed as 
$$T(\ell) = \text{ML}(0.9, 2000 * k^{-1/0.9}).$$ 
A QQ plot (Figure \ref{fig:bitcoin-diag}) of the exceedance times at the 
level of the 200th 
largest event lends credence to a heavy-tailed distribution with tail
parameter near the value $1$.

## Solar Flare Data

```{r solar-flare-data, fig.cap="\\label{fig:solar-flare-data}The solar flare time series data."}
flares_ctrm <- ctrm(flares)
plot(flares_ctrm, p = 0.01)
```

The "complete Hard X Ray Burst Spectrometer event list" 
\footnote{\url{https://umbra.nascom.nasa.gov/smm/hxrbs.html}} 
@HXRBS is a 
comprehensive reference for all measurements of the Hard X 
Ray Burst Spectrometer on NASA's Solar Maximum Mission from the time of 
launch on Feb 14, 1980 to the end of the mission in Dec 1989. 12,776 
events were detected, with the "vast 
majority being solar flares". The list includes the start time, peak time, 
duration, and peak rate of each event. We have used "start time" as the 
variable for event times, and "peak rate" as the variable for event 
magnitudes (Figure \ref{fig:solar-flare-data}).

Figure \ref{fig:flares} shows the fitted tail and scale parameters 
of the Mittag-Leffler distribution. The dotted line for the tail 
parameter is at $\beta = 0.85$, yielding the plot of (transformed) scale 
estimates $\hat \sigma_0 = k^{1/0.85} \hat \sigma(\ell)$. 
We suggest the estimate $\sigma_0 = 2.5 \times 10^7$ seconds 
$\approx 289$ days. 
The exceedance times of a threshold $\ell$ as high as the $k$-measurement 
are hence Mittag-Leffler distributed as
$$ T(\ell) \sim {\rm ML}(0.85, 289  k^{-1/0.85} \text{ days}).$$

# Discussion

Comparing plots of the simulated data in Figure \ref{fig:thresholdedBursty}
with plots of the 
empirical data \ref{fig:bitcoin-data} and \ref{fig:solar-flare-data}, 
it can be seen that the empirical data show long time intervals without 
*any* event occurring, whereas for real-world data events appear to happen 
continuously. 
This seeming discrepancy vanishes, however, when all but the $k_\text{max}$
largest observations are discarded. 
We repeat that the CTRM model focuses on extreme events and interprets 
smaller events as noise.

The CTRM model, moreover, assumes a "pure" power-law for the event 
inter-arrival times. 
Real datasets, however, are often exponentially 
tapered or truncated, see e.g. the discussion of tempered power-law 
distributions by @MeerschaertRoyQin and the truncated Pareto distribution 
by @Aban06. 
This means that the fit to a power-law
distribution may seem adequate for medium-size time scales, 
but that at very large time-scales the power-law character of the 
distribution can weaken or disappear.
This behaviour seems to appear in both the Bitcoin data and the Solar-Flare
data: the Hill plot for the tail parameter is increasing towards lighter
tails as the threshold reaches very high values. 
This may mean that the CTRM model may overestimate the heaviness of the
tail of the waiting time distribution; or that a CTRM base on a *tempered* 
power-law renewal process might be a more realistic model.

# Conclusion

We have proposed the CTRM model for the extremes of "bursty" events. 
Burstiness is a phenomenon of intermittency which has received a lot of 
attention in the recent statistical physics literature. 
The CTRM model is a straightforward generalization of the POT (Peaks over 
Threshold) model, for which the inter-arrival times between threshold 
crossings are exponentially distributed. 
If the Mittag-Leffler tail parameter of the CTRM equals $1$, then the 
CTRM model reproduces this behaviour exactly, showing that the CTRM model 
nests the POT model.

The scale parameter $\sigma$ of the inter-event distribution follows a 
power-law $\sigma \propto p^{-1/\beta}$, where $\beta \in (0,1)$ is the 
tail parameter of the inter-event distribution and $p$ the probability 
that an event crosses a given threshold. 
From this power law, we have constructed "stability plots", from which 
estimates of $\beta$ and $\sigma$ can be read off.

We have demonstrated the applicability of the CTRM model with two 
real-world datasets. 
Although the CTRM model possibly overestimates the heaviness of the tail 
of the inter-arrival distribution, it clearly captures that 
the inter-arrival times of threshold crossings scale non-linearly with 
the treshold crossing probability, and follow a clearly non-Poissonian 
behaviour. 
Thus we have shown that the CTRM model is a useful tool for the modelling 
of extremes of bursty events, which requires very little computation and 
which allows for straightforward ways of checking and criticizing model 
assumptions. 
The analyses from this paper are readily reproduced via the R package 
`CTRM`, available at \url{https://github.com/strakaps/CTRM}.

## Acknowledgements

The authors would like to thank Peter Scheffler for insights on 
stochastic process limits for CTRMs, and Gurtek Gill who helped create
the MittagLeffleR R-package.

For our inference on the CTRM model, we extract realizations of 
$X(\ell)$ and $T(\ell)$ from time series data by considering thresholds 
$\ell$ ranging over some interval $[\ell_0, x_R)$. 
(A reasonable default value for $\ell_0$ is the $95$-th
percentile of $J$.)
Write $X(\ell,i)$ for the $i$-th exceedance of level $\ell$, 
and $T(\ell,i)$ for the inter-arrival time between
$X(\ell,i-1)$ and $X(\ell,i)$. 
Thus for any threshold $\ell \in [\ell_0, x_R)$,
we construct sequences $\{T(\ell,i)\}_{i}$ and 
$\{X(\ell,i)\}_{i}$. 
These sequences are i.i.d. because a renewal occurs at each threshold
crossing time, and independent because $W_k$ and $J_k$ are uncoupled.
Extracting the exceedances $X(\ell)$ for inference is a standard procedure
known as the Peaks Over Threshold (POT) approach in extreme value theory
[@beirlantBook].
Figure \ref{fig:thresholdedBursty} illustrates the exceedances 
$X(\ell, i)$ in red and the times until exceedance $T(\ell, i)$ as the 
preceding intervals marked by blue crosses.

A standard result from extreme value theory provides the asymptotic 
distribution of $X(\ell)$ as $\ell \uparrow x_R$. For later use, we also 
include the scaling limit of the cumulative maximum $M(n)$.

**Theorem GEV** (e.g. Chapter 2 & Section 5.3, @beirlantBook):

: Suppose the distribution of $J_k$ is continuous. Then there
exist non-decreasing norming functions $a(c)$ and $d(c)$ such that
as $c \to \infty$, the running maximum\footnote{We extend $M(n)$ 
continuously to $[0, +\infty)$ via $M(c) := M(\lfloor c \rfloor)$ where 
$\lfloor c \rfloor$ is the 
greatest integer not greater than $c$.} 
$M(c)$ converges weakly 
to a *Generalized Extreme Value Distribution* with shape
parameter
$\xi$:
\begin{align}
[M(c) - d(c)] / a(c) \stackrel{d}{\to} A, 
\quad
\mathbf P(A \le z) = G(z | \xi) = \exp\left(-[1+\xi z]^{-1/\xi}\right),\label{eq:GEV}
\end{align}
defined for all $z$ such that $1 + \xi z > 0$. 
This definition extends continuously to $\xi = 0$ via 
$$G(z | 0) = \exp(-\exp(-x)).$$
For $\mu$ and $\sigma > 0$, we write ${\rm GEV}(\xi, \mu, \sigma)$ for the 
probability distribution of the random variable $\sigma A + \mu$.

    Moreover, asymptotically as $\ell \uparrow x_R$,
$$\mathbf P(J - \ell > y | J > \ell) 
\sim 1 - H(y),$$
where 
$$H(y) = \begin{cases}
1 - (1+ \xi y / \tilde \sigma)^{-1/\xi}, & y \in (0,\infty) \text{ if } \xi > 0,\\
1 - \exp(-y/\tilde \sigma), & y \in (0,\infty) \text{ if } \xi = 0,\\
1 - (1+ \xi y / \tilde \sigma)^{-1/\xi}, & y \in (0, -\sigma / \xi) \text{ if } \xi < 0
\end{cases}$$

    and where $\tilde \sigma := \sigma + \xi(\ell-\mu)$.
The CDF $H(y)$ is said to belong to the *Generalised Pareto family*
$GP(\xi,\bar \sigma)$.

# CTRM Scaling limits

## Waiting times

We will assume that the waiting times $W_k$ have infinite mean,
i.e. a tail parameter $\beta \in (0,1)$ and regularly varying tail probabilities:
$\mathbf P(W > t) \sim L(t) t^{-\beta}$ as $t \uparrow \infty$ for
some slowly varying function $L(t)$. As usually, $f(t) \sim g(t)$
means that the quotient of $f(t)$ and $g(t)$ converges to $1$. The
law of the $W$ then lies in the domain of attraction of a stable law,
in the sense that the distributional limit 
\begin{align} \label{eq:Wscale}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
exists, for a regularly varying scaling function
\begin{equation} \label{eq:bScale}
b \in {\rm RV}_\infty(1/\beta)
\end{equation}
[See e.g. Section 2.9.2 in @beirlantBook for a short introduction to 
regular variation]. 
The limit is then a positively skewed stable distribution, 
whose scale parameter is $1$ if $b(n)$ is chosen accordingly
[see e.g. Section 3.7 in @MeerschaertSikorskii]. 
This distribution is most concisely defined via its Laplace transform
\begin{align} \label{theD}
\mathbf E[\exp(-sD)] = \exp(-s^\beta).
\end{align}
Moreover, the following functional limit theorem holds 
[see e.g. Remark 4.17 in @MeerschaertSikorskii]: 
\[
(W_1 + \ldots + W_{\lfloor ct \rfloor})/b(c) \overset{d}{\longrightarrow} D(t), \quad c \to \infty
\] 
with convergence in the Skorokhod $J_1$ topology. The limit
$D(t)$ is a stable subordinator, i.e. an increasing Levy process with
Laplace transform $\exp(-t s^\beta)$.

It is well known [see e.g. @limitCTRW] that the renewal
process then satisfies the functional limit
$$
N(ct)/ \tilde b(c) \stackrel{d}{\to} E(t) = \inf\{r: D(r) > t\}, 
\quad c \to \infty
$$
for a scaling function $\tilde b(c) \in {\rm RV}_\infty(\beta)$ which is asymptotically inverse to $b(c)$, in the sense of @seneta (p.20): 
$$
b(\tilde b(c)) \sim c \sim \tilde b(b(c)).
$$
[also see Prop 4.15 in @MeerschaertSikorskii]. 
The limit process $E(t)$ is called the _inverse stable subordinator_ [@invSubord].

## Magnitudes and Extreme Value Distributions

The extremal limit theorem then allows for an extension to functional 
(i.e. pathwise) limits: assume that the norming sequences $a(c)$ and 
$d(c)$ are as in Theorem GEV. Then 
$$
[M(ct) - d(c)] / a(c) \stackrel{d}{\to} A(t),
\quad c \to \infty.
$$
Here convergence holds weakly in Skorokhod's $J_1$ topology, and the limit
process $A(t)$ is an _extremal process_, with
finite-dimensional distributions given by
\begin{align}
\label{eq:fin-dim-A}
\mathbf P(A(t_i)\leq x_i,1\leq i \leq d) 
= F_A(\wedge_{i=1}^d x_i)^{t_1} 
F_A(\wedge_{i=2}^d x_i)^{t_2-t_1}
\ldots F_A(x_d)^{t_d-t_{d-1}},
\end{align}
where $F_A(x)$ is a GEV distribution [see e.g. @resnick2013extreme].

## CTRM limit

The CTRM $V(t)$ results from the running maximum $M(n)$ via a time change by the renewal process $n = N(t)$. This is also reflected in its corresponding limit process:

**Theorem MS1** [@MeerschaertStoev08]:

: The CTRM process $V(t) = M(N(t))$ satisfies the following functional scaling limit in the Skorokhod $J_1$ topology: 
$$
[V(ct) - d(\tilde b(c))] / a(\tilde b(c)) \stackrel{d}{\to} 
A(E(t)), \quad c \to \infty.
$$

The distribution of the _hitting time_ of a level $\ell^*$ by the limit 
process $A(E(t))$ is also known, and we give it below for the convenience 
of the reader:

**Theorem MS2** [@MeerschaertStoev08]:

: Let $F_A$ be the CDF of a GEV distribution, and let $A(t)$ be the
extremal process corresponding to $F_A$ via \eqref{eq:fin-dim-A}. 
For a given threshold $\ell^*$ in the support of $F_A$,
the hitting time
$$
B(\ell^*) = \inf\{t: A(E(t)) > \ell^*\} 
$$
is equal in distribution to $(-\log F_A(\ell^*))^{-1/\beta} X^{1/\beta} D$,
where $X \sim {\rm Exp}(1)$ and $D$ (defined in \eqref{theD}) are 
independent.

*Proof:*
Let the scaling functions $a(c)$ and $d(c)$ be as in Section 
\ref{magnitudes-and-extreme-value-distributions}, and assume WLOG 
$\ell = a(n) \ell^* + d(n)$ for some arbitrary but fixed $\ell^*$, out of the support of $A$. 
We first note that due to \eqref{eq:GEV}
\begin{equation} 
F_J (a(c) \ell^* + d(c))^c \stackrel{d}{\to} F_A(\ell^*), \text{ as } c \to \infty.
\end{equation}
\noindent Taking the logarithm on each side and using the relationship  $z-1 \sim \log(z)$ as $z \to 1$, we get
\begin{equation}
c (1-F_J(a(c) \ell^*+d(c))) \stackrel{d}{\to} - \log F_A(\ell^*).
\end{equation}
\noindent Hence we have 
\begin{equation}
c \cdot p \stackrel{d}{\to} - \log F_A(\ell^*) \label{eq:np_conv}
\end{equation}
\noindent as $\ell \uparrow x_R$.
Now we get

**Remark:**

: By the above theorem, the inter-exceedance times are i.i.d. and
approximately Mittag-Leffler distributed, which means that the exceedance
times form a *fractional Poisson process* [@Laskin2003]. 
Since the Mittag-Leffler distribution nests the exponential distribution 
(the special case where $\beta = 1$), the fractional Poisson process 
generalizes the standard Poisson process. 
As the threshold height increases, the threshold crossing events are 
effectively "thinned out", and the thinned (or rarefied) processes retain
the characteristics of a fractional Poisson process [@Gorenflo2010], 
with a growing scale parameter $b(1/p)$.

**Remark:**

: A fractional Poisson process can be represented as a
standard Poisson process $N(t)$, time-changed by an inverse stable 
subordinator $E(t)$ [@Meerschaert2010b]. 
This pattern can be made sense of in our context: 
The exceedance times for a standard extreme value process $A(t)$ form a 
standard Poisson process $N(t)$. Accordingly, the exceedance times 
of the CTRM limit $A(E(t))$ must be $N(E(t))$, a fractional Poisson 
process.

\newpage

>>>>

  This article generalizes the Peaks Over Threshold (POT) model to the setting 
  where inter-event times are heavy-tailed and i.i.d.. For high thresholds and 
  infinite-mean waiting times, we show that the times between threshold 
  crossings are Mittag-Leffler distributed, and thus form a "fractional 
  Poisson Process" which generalizes the standard Poisson Process.
  We provide graphical means of estimating model parameters and assessing 
  model fit. Along the way, we apply our inference method to a real-world
  bursty time series, and 
  show how the Mittag-Leffler distribution has a memory which affects 
  the predictive distribution of the next extreme event.

@@

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Section 3.
We give a statistical procedure to estimate model parameters via stability plots 
in Section 5, but to set the stage we first discuss inference for the 
Mittag-Leffler distribution in Section 4. 
Diagnostic plots for model criticism are discussed in Section 6. 
In Section 7, we discuss the non-memorylessness property of the 
Mittag-Leffler distribution, and how it affects the predictive distribution 
for the next threshold crossing event. 
Section 8 concludes.
All statistical computations have been done in R, and all code, data, as 
well as the source code for this manuscript has been organized into an R 
software package, available at <https://github.com/strakaps/CTRM>.

@@

: Let the waiting times $J_k$ be in the domain of attraction of a 
positively skewed sum-stable law with stability parameter $0 < \beta < 1$; 
more precisely,
\begin{align} \label{eq:stability}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
for a function $b(n)$ which is regularly varying at $\infty$ with 
parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$. 
Write $p := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p)} \to W_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $W_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sW_\beta)] = \frac{1}{1+s^\beta}.
$$

@@

**Remark:**
: If $\beta = 1$, the result of the Theorem above is standard, see e.g. 
Equation (2.2) in @Gut1999.

@@

```{r solar-flare-tail-scale, message=FALSE, fig.height=4.5, fig.width=8, fig.cap="\\label{fig:flares}Stability plots for the tail and scale parameter of the Mittag-Leffler distribution of the Solar Flare dataset. Dotted horizontal lines are at $\\beta = 0.85$ and $\\sigma_0 = 3 \\times 10^7$ seconds $\\approx 0.95$ years."}
thin_flares_ctrm <- flares %>% ctrm() %>% thin(k = 700)
par(mfrow=c(1,2))
MLestimates(thin_flares_ctrm, tail = 0.9, scale = 3 * 1E7)
```

@@

# Discussion & Conclusion

We have extended the POT (Peaks over Threshold) model, a mainstay of 
extreme value theory, to "bursty" time series, which have been studied 
intensively in statistical physics. 
Burstiness is characterized by power-law waiting times between events, 
and we have shown that the Mittag-Leffler distribution arises naturally 
as a scaling limit for the inter-exceedance times of high thresholds. 
Moreover, we have derived the following non-linear scaling behaviour: 
$\sigma \sim p^{-1/\beta}$, 
where $\sigma$ is the scale parameter of the distribution of threshold 
exceedance times, 
$p$ is the fraction of magnitudes above the threshold, 
and $\beta$ the exponent of the power law.

The "anomalous" scaling behaviour in the bursty setting entails two 
phenomena: 
i) a heavy tail of the interarrival time distribution of 
threshold crossings (long rests), and 
ii) a high propensity for more threshold crossing events immediately after 
each threshold crossing event (bursts). 
The Mittag-Leffler distribution captures both phenomena, due to its heavy
tail as well as its streteched exponential (peaked) asymptotics for small 
times. 
It generalizes the exponential distribution, and 
in the solar flare data example, this generalization 
is warranted, because the likelihood-ratio test is strongly significant.

When we introduced the MRP model, we assumed that all events are i.i.d. 
This assumption is likely sufficient but not necessary for our limit theorem to hold. 
Morevoer, any data below a (minimum) threshold $\ell_0$ is discarded in 
our inference procedure, and hence need not satisfy the i.i.d. assumption.
For the purposes of statistical inference, we merely require that the 
inter-threshold-crossing times are i.i.d.

For some bursty time series, the current MRP model may be too rigid, 
because often the heavy-tailed character of the inter-arrival time
distribution does not hold at all time scales; rather, it often applies at 
short and intermediate time scales, and is truncated or tempered 
(reverting to an exponential distribution) at 
very long time scales, see e.g. @MeerschaertRoyQin and @Aban06. 
In such situations, a "tempered" Mittag-Leffler distribution may provide 
a more realistic fit, which we aim to introduce in follow-up work.

## Acknowledgements {-}

The authors would like to thank Prof. Peter Scheffler for insights on 
stochastic process limits for CTRMs, and Gurtek Gill who helped create
the MittagLeffleR R-package.

# Validating our inference method on simulated data

To test our inference method via stability plots, we have simulated `r params$n` 
independent
waiting time and magnitude pairs $(W_k, J_k)$ (see upper panel in 
Figure \ref{fig:thresholdedBursty}). In order to have exact 
analytical values available for $\beta$ and $\sigma_0$, a distribution for 
$W_k$ needs to be chosen for which $b(n)$ from \eqref{eq:stability} is known. 
If we choose $W_k \stackrel{d}{=} D$, where $D$ is as in \eqref{eq:stability}, 
then due to the stability property we have the *equality* of distribution
$W_1 + \ldots + W_n \stackrel{d}{=} b(n) D$, 
for $b(n) = n^{1/\beta}$. 
Using the parametrisation of @SamorodnitskyTaqqu, a few lines of 
calculation (see e.g. the vignette on parametrisation in @MittagLeffleR) 
show that $D$ must have the stable distribution 
$S_\beta(\cos(\pi \beta/2)^{1/\beta}, +1, 0)$, which is 
implemented in the R package `stabledist` by @stabledist.

@@

By the Theorem, the distribution 
of $T(\ell)$ is approximately 
$$
{\rm ML}(\beta, p^{-1/\beta}) 
= {\rm ML}(\beta, k^{-1/\beta} n^{1/\beta}),
$$
which means 
$\sigma_0 = n^{1/\beta}$.
The distribution of $J_k$ is irrelevant for the inference on $\beta$ and 
$\sigma_0$ (we have chosen unit exponential random variables). 
Figure \ref{fig:sim} displays plots of $\hat \beta(\ell)$ and 
$\hat \sigma(\ell)$ vs. $k$; recall that $k$ is the index of the order
statistics of $J_k$ at which the threshold $\ell$ is placed. 
Dotted lines show 95% confidence intervals, which are derived from the 
asymptotic normality of the log-moments estimators [@Cahoy2013] and the $\delta$-method [@MittagLeffleR].
The dashed lines show the actual values of $\beta$ resp.\ $\sigma_0$, 
showing that our inference method identifies the 
parameters correctly.

# References {-}
